---
title: "Class 10: Scientific Publishing and the Replication Crisis"
bibliography: ../references.bib
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
```

## Slides

The slides are available online as an HTML file. You can also download them in a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/10-slides")
```

## Readings

I don't expect you to read all these, but in case you're curious, here are some papers mentioned in class:

-   @opensciencecollaborationEstimatingReproducibilityPsychological2015

-   @simmonsFalsePositivePsychologyUndisclosed2011

-   @gouldSameDataDifferent2025

-   @turnerSelectivePublicationAntidepressant2008

For a refresher on hypothesis testing and statistical power, you could can have a look at the begging of [chapter 2](https://lakens.github.io/statistical_inferences/02-errorcontrol.html) in @lakensImprovingYourStatistical2022d

Since the lecture did not really cover this, if you want to know what's wrong with scientific publishing, [this old (but gold!) article of the Guardian](https://www.theguardian.com/science/2017/jun/27/profitable-business-scientific-publishing-bad-for-science) is a good read. It's also available as a podcast.

## Data

-   [{{< fa table >}} `imaginary_beatle_experiment.csv`](../data/imaginary_beatle_experiment.csv)

## Assignment

No assignment for this session `r emoji::emoji("tada")`
