---
title: "Class 12: Meta Analyses"
bibliography: ../references.bib
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
```

## Slides

The slides are available online as an HTML file. You can also download them in a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/12-slides")
```

## Readings

In case you are curious to learn more about meta-analysis, check out this great (and free) book:

-   [@harrerDoingMetaAnalysisHandsOn2021](https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/)

Other things:

- [Interpreting Cohen's d Effect Size by Kristoffer Magnusson](https://rpsychologist.com/cohend/)
- the [PRISMA guidelines](https://www.prisma-statement.org/) for systematic reviews and meta-analyses

## Assignment

No assignment for this session `r emoji::emoji("tada")`
