---
title: "Class 13: How to Read and Present a Scientific Paper"
bibliography: ../references.bib
---

```{r}
#| include: false
library(tidyverse)
library(kableExtra)
```


There are no slides for this class. You will need to work on your group presentations (more information to come)

## Readings

For your presentations, you need to read the papers you have been assigne (see next section). You can find PDFs for all papers on Moodle.

## Assignment

For this week's assignment, you have been randomly assigned to read one of 5 papers (see below). Your assignment is to prepare group a presentation on your paper. 

```{r}
#| code-fold: true
names <- c(
  "Jiwoo", "Arthur", "Justin", "Tom", "Alice C.", "Agathe", "Lisa", "Jeanne",
  "Valentine", "Micaela", "Assya", "Cyriac", "Alice G.", "Thelma", "Titouan",
  "Maho", "Selma", "Lucas", "Olga", "Eloïse", "Josephine", "Rivka", "Damian",
  "Candice", "Marie", "Mael", "Ariane", "Camille", "Etienne", "Anitha"
)

# define groups
texts <- tribble(~group, ~text,
                  1, "Simonsohn, U., Simmons, J., & Nelson, L. D. (2022). \n Above averaging in literature reviews. \n + [105] Meaningless Means #1: The Average Effect of Nudging Is d = .43. 
", 
                  2, "Mertens, S., Herberz, M., Hahnel, U. J. J., & Brosch, T. (2022). \n The effectiveness of nudging.
", 
                  3, "dos Santos, Q., Perez-Cueto, F. J. A., Rodrigues, V. M., Appleton, K., \n Giboreau, A., Saulais, L., Monteleone, E., Dinnella, C., Brugarolas, M., & Hartwell, H. (2020). \n Impact of a nudging intervention and factors associated with vegetable dish choice.
", 
                  4, "Bacon, L., & Krpan, D. (2018). \n (Not) Eating for the environment. 
", 
                  5, "Tavernier, R., & Adam, E. K. (2017). \n Text message intervention improves objective sleep hours among adolescents.
", 
                  6, "Diliberti, N., Bordi, P. L., Conklin, M. T., Roe, L. S., & Rolls, B. J. (2004). \n Increased Portion Size Leads to Increased Energy Intake in a Restaurant Meal. 
")


# build data frame to which names can be assigned
groups_empty <- crossing(texts, spot_by_group = 1:5) %>% 
  mutate(row_number = 1:nrow(.))

# random group allocation

## make reproducible
set.seed(2025)

# randomly shuffle order of names 
names_random_order <- tibble(name = sample(names, length(names))) %>% 
  mutate(row_number = 1:nrow(.))

# pair with groups
random_groups <- right_join(groups_empty, names_random_order, by = "row_number") %>% 
  select(group, name, text, -c(spot_by_group, row_number))

random_groups %>%
  kable(escape = FALSE, col.names = c("Group", "Name", "Text")) %>%
  kable_styling(full_width = FALSE, position = "left", font_size = 11)
```
















