---
title: "Class 13: How to Read and Present a Scientific Paper"
bibliography: ../references.bib
---

```{r}
#| include: false
library(tidyverse)
library(kableExtra)
```

There are no slides for this class. This is a practical sessions during which you will work work on your group presentations. 

## Group presentations

### The rules

- You have 5 minutes per group. 
- Followed by 10 mins plenary discussion. 
- Each group member presents (~1min per member).

### Strategy

5mins is short. It requires making a selection of what to tell, and that is difficult. The main goal is to: 

- gather interest among your audience 
- sketch out the main idea

Everything else will be raised in the discussion. You can prepare some back up slides if you anticipate certain questions or plan raising certain discussion points.

### A (possible) outline

1. Why should you care? (~1min)
  - State the research question, make it relatable to the audience, potentially add some context to raise a broader puzzle that is interesting to someone who is not an expert

2. What did they do? (~2mins)
  - Describe the experimental design and process
  - A good technique is to take the perspective of a participant (i.e. describe what they saw, exactly)

3. What do they find? (~1min)
  - Focus on the one main result & authors' conclusions from it
  
4. What’s your take on that? (~1min)
  - You can make the link to the initial puzzle or highlight what they did well / what you found not convincing

## Readings

For your presentations, you have been randomly assigned to read one of 5 papers (see below). You can find PDFs for all papers on Moodle.

```{r}
#| code-fold: true
names <- c(
  "Jiwoo", "Arthur", "Justin", "Tom", "Alice C.", "Agathe", "Lisa", "Jeanne",
  "Valentine", "Micaela", "Assya", "Cyriac", "Alice G.", "Thelma", "Titouan",
  "Maho", "Selma", "Lucas", "Olga", "Eloïse", "Josephine", "Rivka", "Damian",
  "Candice", "Marie", "Mael", "Ariane", "Camille", "Etienne", "Anitha"
)

# define groups
texts <- tribble(~group, ~text,
                  1, "Simonsohn, U., Simmons, J., & Nelson, L. D. (2022). \n Above averaging in literature reviews. \n + [105] Meaningless Means #1: The Average Effect of Nudging Is d = .43. 
", 
                  2, "Mertens, S., Herberz, M., Hahnel, U. J. J., & Brosch, T. (2022). \n The effectiveness of nudging.
", 
                  3, "dos Santos, Q., Perez-Cueto, F. J. A., Rodrigues, V. M., Appleton, K., \n Giboreau, A., Saulais, L., Monteleone, E., Dinnella, C., Brugarolas, M., & Hartwell, H. (2020). \n Impact of a nudging intervention and factors associated with vegetable dish choice.
", 
                  4, "Bacon, L., & Krpan, D. (2018). \n (Not) Eating for the environment. 
", 
                  5, "Tavernier, R., & Adam, E. K. (2017). \n Text message intervention improves objective sleep hours among adolescents.
", 
                  6, "Diliberti, N., Bordi, P. L., Conklin, M. T., Roe, L. S., & Rolls, B. J. (2004). \n Increased Portion Size Leads to Increased Energy Intake in a Restaurant Meal. 
")


# build data frame to which names can be assigned
groups_empty <- crossing(texts, spot_by_group = 1:5) %>% 
  mutate(row_number = 1:nrow(.))

# random group allocation

## make reproducible
set.seed(2025)

# randomly shuffle order of names 
names_random_order <- tibble(name = sample(names, length(names))) %>% 
  mutate(row_number = 1:nrow(.))

# pair with groups
random_groups <- right_join(groups_empty, names_random_order, by = "row_number") %>% 
  select(group, name, text, -c(spot_by_group, row_number))

random_groups %>%
  kable(escape = FALSE, col.names = c("Group", "Name", "Text")) %>%
  kable_styling(full_width = FALSE, position = "left", font_size = 11)
```

## Assignment

If you do not manage to finish your presentations in class, you will need to finish them at home before next session. 

















