---
title: "Problem set 1"
author: "Put your name here"
format: 
  html:
    toc: true
  pdf:
    toc: true
  docx:
    toc: true
---

```{r packages}
library(tidyverse)
```

# Learning R

# Read the data

> Read the `cars.csv` data into R. Make sure to use the correct path ("data/cars.csv"). Name the data frame "cars" when reading it in. You don't need to understand what all the variables mean. 

```{r}
cars <- read_csv("../data/cars.csv")
```

# What's the class of the `model` and the `year` variable?

```{r}
class(cars$model)
class(cars$year)
```

> Subset the cars data by selecting only rows that correspond to the manufacturer "honda" and that shows only the columns for models and the year. Name that subset "honda_data" and print it. 

```{r}
honda_data <- cars[cars$manufacturer == "honda", c("model", "year")]

# alternative
honda_data <- cars %>%
  filter(manufacturer == "honda") %>%
  select(model, year)
```

# My first plots

> You haven't learned about plots yet. But to give you a taste for what's coming, execute the code chunk below and let the magic happen. Make sure your data frame is named "cars" for this to work

A plot on the distance that cars can travel per gallon. Note that we will hide the code when rendering by setting `echo: false`.

```{r plot-data}
#| echo: false
ggplot(cars, aes(x = hwy)) +
geom_histogram() +
labs(
title = "A Histogram",
x = "Higway MPG*", 
caption = "*miles per gallon, is the distance, measured in miles, that a car can travel per gallon of fuel."
)
```